\chapter{Model}
\paragraph{What model was used}
The model used for this research was a Generalized Additive Mixed Model (GAMM).
To explain what a GAMM is, we first have to introduce some lower-level models.\\
\subsection{From LM to GLM to GAM to GAMM}
\subsubsection{LM}
Linear regression is a method used to describe how the average 
value of a numerical outcome changes according to a linear 
combination of predictors.
This kind of regression can be used to represent relationships 
between variables.
However, the linear regression model relies on several strong assumptions, 
in particular that the response variable is continuous, 
normally distributed, and linearly related to the predictors. 
In many real-world applications, these assumptions are not appropriate. 
As in our case our outcome is non-normally distributed, we need
to use a different model.
\subsubsection{Linear Model (LM)}
Test formula: $x = 1$
\begin{equation}
Y_i = \beta_0 + \beta_1x_{i1} + \cdots + \beta_px_{ip} + \varepsilon_i,
\end{equation}

where

\begin{equation}
\varepsilon_i \sim N(0,\sigma^2).
\end{equation}

and:
\begin{itemize}
    \item $Y_i$ is the response variable;
    \item $\beta_0$ is the intercept;
    \item $\beta_1,\ldots,\beta_p$ are the regression coefficients associated with the predictors;
    \item $x_{i1},\ldots,x_{ip}$ are the values of the predictors for observation $i$;
    \item $\varepsilon_i$ is the random error term, assumed to be independently and identically distributed as
    \[
    \varepsilon_i \sim N(0,\sigma^2).
    \]
\end{itemize}

\subsubsection{From LM to GLM}
To address these above-mentioned limitations, we can recurr to 
the Generalized Linear Model (GLM), which extends the classical
linear model by allowing the response variable to follow 
distributions from the exponential family and by introducing a 
link function that connects the expected value of the response 
to the linear predictor.
\subsubsection{Generalized Linear Model (GLM)}

\begin{equation}
g(\mu_i)=\eta_i=
\beta_0+\beta_1x_{i1}+\cdots+\beta_px_{ip},
\end{equation}

where

\begin{equation}
\mu_i=E(Y_i),
\end{equation}

and:
\begin{itemize}
    \item $\mu_i=E(Y_i)$ is the expected value of the response variable;
    \item $g(\cdot)$ is the link function;
    \item $\eta_i$ denotes the linear predictor.
\end{itemize}

\subsubsection{From GLM to GAM}
While GLMs provide a flexible framework for handling different
types of response variables, they still assume a linear 
relationship between the transformed expected value of the
response and the predictors. In practice, this assumption would
still be too restrictive, since many relationships in real 
data are non-linear.\\
To overcome this limitation, Generalized Additive Models (GAMs) 
extend GLMs by allowing the linear predictor to include smooth, 
functions of the covariates. This makes it possible to model 
complex and potentially non-linear effects of predictors on the 
response variable while maintaining interpretability.
\subsubsection{Generalized Additive Model (GAM)}

\begin{equation}
g(\mu_i)=
\beta_0+
f_1(x_{i1})+
f_2(x_{i2})+
\cdots+
f_p(x_{ip}) +
\varepsilon_i,
\end{equation}

where $f_j(\cdot)$ are smooth functions estimated from the data.

\subsubsection{From GAM to GAMM}
Although GAMs allow for flexible functional forms, they still 
assume that observations are independent. However, in our case 
(where we hierarchically grouped the observations getting Continents --> Countries) 
this assumption is violated.\\
To account for this dependence structure, Generalized Additive 
Mixed Models (GAMMs) further extend GAMs by incorporating random
effects into the model. This allows the model to capture both 
non-linear relationships through smooth functions and correlation
structures through random effects, making GAMMs particularly 
suitable for hierarchical data.

\subsubsection{Generalized Additive Mixed Model (GAMM)}

\begin{equation}
g(\mu_i)=
\beta_0+
\sum_{j=1}^{p}f_j(x_{ij})+
\mathbf{Z}_i\mathbf{b} +
\varepsilon_i,
\end{equation}

where

\begin{equation}
\mathbf{b}\sim N(\mathbf{0},\mathbf{D})
\end{equation}

represents the random effects, and:
\begin{itemize}
    \item $\mathbf{b}$ is the vector of random effects;
    \item $\mathbf{Z}_i$ is the design matrix associated with the random effects;
    \item $\mathbf{b}\sim N(\mathbf{0},\mathbf{D})$.
\end{itemize}

Several candidate models were compared, and the best-fitting 
model was selected according to the Akaike Information Criterion (AIC).\\
According to AIC, in fact, our final model had the lowest value 
compared to the other prototypes, and during this process, some 
variables have been cut out in order to guarantee a rigorous but 
efficient selecting criteria requiring a certain amount of 
statistical significance for a variable to be included.\\
This has led to a model composed by: the Stringency Index as a 
linear term, smooth functions for GNI per Capita and Internet 
Users, a tensor interaction between Year and Depression Rates, 
and random intercepts for countries nested within continents.

\begin{align}
\text{Sales}_{ijt} = & \ \beta_0 + \beta_1 \text{StringencyIndex}_{ijt} \\
& + f_1(\text{GNIpc}_{ijt}) \\
& + f_2(\text{InternetUsers}_{ijt}) \\
& + f_3(\text{Year}_{ijt}, \text{DepressionRate}_{ijt}) \\
& + u_j + v_i + \varepsilon_{ijt}
\end{align}

where

\begin{itemize}
    \item $i$ indexes countries;
    \item $j$ indexes continents;
    \item $t$ indexes years;
    \item $\beta_0$ is the intercept;
    \item $\beta_1$ represents the linear effect of the Stringency Index;
    \item $f_1(\cdot)$ and $f_2(\cdot)$ denote smooth functions capturing the potentially non-linear effects of GNI per capita and Internet Users;
    \item $f_3(\cdot,\cdot)$ represents a tensor-product smooth interaction between Year and Depression Rates;
    \item $u_j$ is the random intercept associated with continent $j$;
    \item $v_{ij}$ is the random intercept associated with country $i$, nested within continent $j$;
    \item $\varepsilon_{ijt}$ is the error term.
\end{itemize}

\paragraph{Model Results}
In this section, empirical results of the GAMM estimation, 
focusing on both parametric and non-linear effects
will be presented.

\begin{table}[ht]
\centering
\caption{GAMM Estimation Results}
\begin{tabular}{l|c|c|c|c}
\hline
\textbf{Variable} & \textbf{Estimate / edf} & \textbf{Std. Error / Ref.df} & \textbf{Statistic} & \textbf{p-value} \\
\hline

\multicolumn{5}{l}{\textit{Parametric coefficients}} \\
Intercept & 6.541 & 0.357 & 18.300 & $<0.001^{***}$ \\
Stringency Index & 0.098 & 0.048 & 2.029 & 0.043$^{*}$ \\

\hline
\multicolumn{5}{l}{\textit{Smooth terms}} \\
s(GNI per Capita) & 2.725 & 2.725 & 32.45 & $<0.001^{***}$ \\
s(Internet Users) & 3.453 & 3.453 & 20.41 & $<0.001^{***}$ \\
te(Year, Depression Rates) & 7.489 & 7.489 & 66.22 & $<0.001^{***}$ \\

\hline
\end{tabular}
\vspace{2mm}
\begin{flushleft}
\footnotesize
$^{***} p < 0.001$, $^{*} p < 0.05$ \\
Adjusted $R^2 = -0.241$, $n = 293$
\end{flushleft}
\end{table}

Note: the number of observations taken in consideration for this
model is lower than the total number of observations present
in the final dataset. This is due to the inconsistency of
some observations' data.

DA MODIFICARE IL MODELLO IN QUANTO AVEVA R^2 NEGATIVO